Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-16 qpip0.tex

In her groundbreaking work, Mahadev~\cite{FOCS:Mahadev18a} gives a $\QPIP_0$ protocol for $\BQP$ languages.
The core of this work is a 4-round $\QPIP_0$ protocol $\PiMeasure$ that lets a $\BQP$ machine ``commit a $XZ$ measurement" to a classical machine.
Intuitively, the verifier chooses a string $h$ specifying  the measurement he wants to make, and send encoded $h$ to the prover. The prover ``commits" to a state $\rho$ with the encoded $h$ and replies. The verifier then uniformly chose between have two options: do a \emph{testing round} or do a \emph{Hadamard round}. He send his choice between the two to the prover have the  prover reply. If the verifier chose testing, he checks the prover's two replies and rejects if he see a inconsistency. If the verifier chose the Hadamard round, he calculate $M_{XZ}(\rho,h)$ based on the replies. Mahadev~\cite{FOCS:Mahadev18a} proves a ``binding" property of the above protocol: if a prover's strategy would always succeeds on the testing round, then that strategy must binds to a $\rho$, meaning that no matter what $h$ the verifier has picked at the beginning, the verifier can get $M_{XZ}(\rho,h)$  with the same $\rho$ if he had chosen the Hadamard round.

@@

Protocol:
\begin{enumerate}
	\item \label{step:measure1} The verifier generates a public and secret key pair $(pk, sk)\leftarrow\cVMeasure{1}(1^\lambda, h)$. It sends $pk$ to the prover.
	\item \label{step:measure2} The prover generates $(y, \sigma)\leftarrow\cPMeasure{2}(pk, \rho)$.
		$y$ is a classical ``commitment", and $\sigma$ is some internal state.
		He sends $y$ to the verifier.
	\item \label{step:measure3} The verifier samples $c\xleftarrow{\$}\zo$ uniformly at random and sends it to the prover. $c=0$ indicates a \emph{testing round}, while $c=1$ indicates a \emph{Hadamard round}.
	\item \label{step:measure4} The prover generates a classical string $a\leftarrow\cPMeasure{4}(pk, c, \sigma)$ and sends it to back to the verifier.
	\item \label{step:output} If it is a testing round ($c=0$), then the verifier generates and outputs $o\leftarrow\cVMeasure{T}(pk, y, a)$ where $o\in\set{\Acc, \Rej}$.
		If it is a Hadamard round ($c=1$), then the verifier generates and outputs $v\leftarrow\cVMeasure{H}(sk, h, y, a)$.
\end{enumerate}
\end{protocol}

@@

\begin{lemma}[binding property of $\PiMeasure$]
\label{lem:urmila-binding}
	Let $\PMeasureStar$ be a $\BQP$ cheating  prover for $\PiMeasure$ and $\lambda$ be the security parameter. Under the QLWE assumption, if $\PMeasureStar$ passes the testing round with probability $1-\negl(\lambda)$, then there exists some $\rho$ so that for all verifier's input $h \in \zo^n$, the verifier's outputs on the Hadamard round is computationally indistinguishable from $M_{XZ}(\rho, h)$.
\iffalse	
	Suppose that for all $\lambda\in\bbN$ and $h\in\zo^*$ \Ethan{or $\zo^n$?},
	$\PMeasureStar$ passes the testing round with probability $1-\negl(\lambda)$.
	Then, under the QLWE assumption, there exists some $\rho$ so that for all $h$,
	the verifier's outputs on the Hadamard round is $\negl(\lambda)$-computationally indistinguishable from $M_{XZ}(\rho, h)$.
\fi
\end{lemma}

@@

There are several problems with using $\PiNaive$ as a $\SampBQP$ protocol. first, since the verifier don't get a sample if he had chosen the testing round in Step~\ref{step:c-urmila-in-naive}, so the protocol has  completeness error at least $1/2$. And since Protocol~\ref{proto:urmila4} does not check anything on the Hadamard round, a cheating prover can give up passing the testing round and breaks the commitment on the Hadamard round. Such cheating prover only has constant $1/2$ probability of being caught, resulting in constant soundness error.

% Note that in $\PiNaive$ the verifier only learns the output in Hadamard round.
% So $\PiNaive$ is not a $\QPIP_0$ with good completeness or soundness.

@@

We now introduce our $\QPIP_0$ protocol $\PiSampZ$ for $\SampBQP$.
It is essentially a $m$-fold parallel repetition of $\PiNaive$.
Instead of having each copies running testing round or Hadamard round uniformly at random, we randomly pick one copy to run Hadamard round to get our samples, and run testing round on all other $m-1$ copies.
In the description of our protocol below, we describe $\PiNaive$ and $\PiMeasure$ in details  in order to introduce notations that we need in our analysis.

@@

Inputs:
	\begin{itemize}
		\item Security parameter $1^\lambda$ for $\lambda\in\bbN$.
		\item Accuracy paramter $1^{1/\eps}$ for the $\SampBQP$ instance
		\item Input $x\in\zo^{\poly(\lambda)}$ for the $\SampBQP$ instance
	\end{itemize}

Ingrediant: Let $m=O(1/\eps^2)$ be the number of parallel repetitions to run.

@@

The intuition behind this theorem is that the we send out $m$ independent copies of the naive $\QPIP_0$ and do testing on randomly chosen $m-1$ copies of them, and do Hadamard round and calculate results from the remaining one copy. Therefore if the prover want to cheat, i.e. send something not binding in the Hadamard round copy, he would be caught with probability $1-1/m$. However, the prover might create correlations between copies that let him cheat with probability much higher than $1/m$. This is a common problem in trying to create parallel repetition protocols. We are able to bound the soundness error as $O(1/\sqrt{m})$. To prove that, we use the following partition lemma from \cite{parallelrep}, which intuitively says that for each $i\in[m]$, there exist two efficient ``projectors" \footnote{Actually they are not projectors, but for the simplicity of this discussion let's assume they are.} $G_{0,i}$ and $G_{1,i}$ in the prover's internal space with $G_{0,i}+G_{1,i} \approx Id$.  $G_{0,i}$ represents the subspace that are rejected with high probability in the testing round of $i$-th copy; $G_{1,i}$  represents the subspace accepted with ``high probability"\footnote{This is also not technically correct as we will discuss later.} in the testing round of $i$-th copy.  $G_{0,i}$ and $G_{1,i}$ need to be efficient because $\PiMeasure$ only gives computational security. Once we have $G_{0,i}$ and $G_{1,i}$, we can split up the prover's internal state $\ket{\psi}$ like conditional probabilistic events.  We  split up  $\ket{\psi}$ into $G_{0,1}\ket{\psi}+ G_{0,2}G_{1,1}\ket{\psi} + G_{0,3}G_{1,2}G_{1,1}\ket{\psi} + \dots +  G_{0,m} G_{1,m-1}\dots G_{1,2}G_{1,1}\ket{\psi} + G_{1,m} G_{1,m-1}\dots G_{1,2}G_{1,1}\ket{\psi}$, where for example, $G_{0,2}G_{1,1}\ket{\psi}$ intuitively corresponds to passing the testing round of the first copy then getting rejected on the testing round of the second copy. Since each of the  substates in  $\{G_{0,1}\ket{\psi}, G_{0,2}G_{1,1}\ket{\psi} , G_{0,3}G_{1,2}G_{1,1}\ket{\psi} , \dots ,  G_{0,m} G_{1,m-1}\dots G_{1,2}G_{1,1}\ket{\psi} \}$ all have  probability $1-1/m$ of being caught by one of the $m-1$ testing rounds, they have well-bounded 
 probability of being accepted in Step~\ref{step:multi-testing} of Protocol~\ref{proto:QPIP0samp}, so the output in  Step~\ref{step:multi-testing} would mostly corresponds to $G_{1,m} G_{1,m-1}\dots G_{1,2}G_{1,1}\ket{\psi}$. However, since $G_{1,i}$ don't commute with $G_{1,j}$, $G_{1,m} G_{1,m-1}\dots G_{1,2}G_{1,1}\ket{\psi}$ is only binding for the $m$-th copy, so this simplified strategy is not quite correct. We were able to fix this issue with a more careful split up of $\ket{\psi}$. Also as a careful reader might have noticed, the prover's space don't happen to split nicely into parts of very high accept probability in the testing round and parts with very low accept probability, as there will always be something in the middle. In \cite{parallelrep} this is solved by doing eigenvalue estimation to calculate the accept probability, then split the space into parts that are accepted with probability higher or lower than a small threshold $\gamma$. However, states with accept probability really close to the threshold $\gamma$ can not be classified, so we need to average over randomly chosen $\gamma$ to have $G_{0,i}+G_{1,i} \approx Id$.

 %but there will always be some states  herefore we introduce  an extra parameter $\gamma$ that corresponds to the probability cut-off line that to distinguish the accepted and rejected part, and we need to average over $\gamma$.

@@

%For example, $G_{0,3} G_{1,5} \ket{\psi}$ corresponds to the events of passing the testing round on fifth copy then fails the the testing round on the third copy.

@@

where $P_{i,pk_i,y_i,acc}$ are projector to the states that $i$-th testing round accepts with $pk_i,y_i$, including the last measurement the prover did before sending $\vec{a}$.  This means that $\ket{\psi_{0,i,\gamma}}$ is rejected by $i$-th testing round with high probability.

@@

% \begin{align*}  
%   \Pr\left[M_{\regX_i}\circ \ext_i\left(\frac{\ket{\{c\}}_{\regC}\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}}{\|\ket{\psi_1}_{\regX,\regZ}\|}\right)\in \Acc_{pk_i,y_i}\right]=1-\negl(\secpar).
%   \end{align*}
This will imply that    $\ket{\psi_{1,i,\gamma}}$ is binding to $i$-th Hadamard round.

@@

where we abuse the notation and use $\ket{\psi_{err,i,\gamma}}$ to denote the error state we get from decomposing $\ket{\psi_{1^{i-1},\gamma}}$. %\hannote{Note that the prover's state is actually a mixed state. In particular, $\ket{\psi}$ depends on $pk$ and we are considering the mixed state averaged over verifier's choice of $pk$.}

@@

We now proceeds to prove that $z_{good}$ is close to the ideal distribution.

@@

The prover tries to run $U_0$ by taking the verifier's input as $i$-th copy of Protocol~\ref{proto:urmila4} in Protocol~\ref{proto:QPIP0samp} and simulating other $m-1$ copies by himself. The prover then picks a uniformly random $\gamma$ and  tries to generate $\ket{\psi_{1^{i-1}1,\gamma}}$ by applying $G_{i,1,\gamma}G_{i-1,1,\gamma} \cdots G_{2,1,\gamma}G_{1,1,\gamma}$. If the prover fails to generate $\ket{\psi_{1^{i-1}1,\gamma}}$, he throws out everything aborts by sending $\bot$ back.   On the fourth round,  If it's a testing round the prover reply with the $i$-th register of $\ext_i\left(\frac{\ket{e_j}_{\regC}\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}}{\|\ket{\psi_1}_{\regX,\regZ}\|}\right)$, where $\ext_i$ is specified in property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}. If it's the Hadamard round  the prover  runs $U$ and check whether every copy except the $i$-th copy would be accepted. If all $m-1$ copies accepts, he reply with the $i$-th copy, otherwise reply $\bot$.

@@

By  property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}, the above strategy is accepted with probability $1-\negl(n)$ when the prover didn't abort. Since the prover's strartegy is also efficient, by Lemma~\ref{lem:naive-qpip0-binding}, 
\begin{align}
    (d,z_{composite,i})\approx_{c, O(\eps)}(d,z_{ideal,i})
\end{align}

>>>>

In her groundbreaking work, Mahadev~\cite{FOCS:Mahadev18a} gives a $\QPIP_0$ protocol for $\BQP$ languages.
The core of this work is a 4-round $\QPIP_0$ protocol $\PiMeasure$ that lets a $\BQP$ machine ``commit a $XZ$ measurement" to a classical machine.
Intuitively, the verifier chooses a string $h$ specifying  the measurement he wants to make, and sends encoded $h$ to the prover. The prover ``commits" to a state $\rho$ with the encoded $h$ and replies. The verifier then uniformly chooses between two options: do a \emph{testing round} or do a \emph{Hadamard round}. He sends his choice between the two to the prover and has the  prover reply. If the verifier chose testing, he checks the prover's two replies and rejects if he sees an inconsistency. If the verifier chose the Hadamard round, he calculates $M_{XZ}(\rho,h)$ based on the replies. Mahadev~\cite{FOCS:Mahadev18a} proves a ``binding" property of the above protocol: if a prover's strategy would always succeeds on the testing round, then that strategy must binds to a $\rho$, meaning that no matter what $h$ the verifier has picked at the beginning, the verifier can get $M_{XZ}(\rho,h)$  with the same $\rho$ if he had chosen the Hadamard round.

@@

Protocol:
\begin{enumerate}
    \item \label{step:measure1} The verifier generates a public and secret key pair $(pk, sk)\leftarrow\cVMeasure{1}(1^\lambda, h)$. It sends $pk$ to the prover.
    \item \label{step:measure2} The prover generates $(y, \sigma)\leftarrow\cPMeasure{2}(pk, \rho)$.
        $y$ is a classical ``commitment", and $\sigma$ is some internal state.
        He sends $y$ to the verifier.
    \item \label{step:measure3} The verifier samples $c\xleftarrow{\$}\zo$ uniformly at random and sends it to the prover. $c=0$ indicates a \emph{testing round}, while $c=1$ indicates a \emph{Hadamard round}.
    \item \label{step:measure4} The prover generates a classical string $a\leftarrow\cPMeasure{4}(pk, c, \sigma)$ and sends it back to the verifier.
    \item \label{step:output} If it is a testing round ($c=0$), then the verifier generates and outputs $o\leftarrow\cVMeasure{T}(pk, y, a)$ where $o\in\set{\Acc, \Rej}$.
        If it is a Hadamard round ($c=1$), then the verifier generates and outputs $v\leftarrow\cVMeasure{H}(sk, h, y, a)$.
\end{enumerate}
\end{protocol}

@@

\begin{lemma}[binding property of $\PiMeasure$]
\label{lem:urmila-binding}
    Let $\PMeasureStar$ be a $\BQP$ cheating  prover for $\PiMeasure$ and $\lambda$ be the security parameter. Under the QLWE assumption, if $\PMeasureStar$ passes the testing round with probability $1-\negl(\lambda)$, then there exists some $\rho$ so that for all verifier's input $h \in \zo^n$, the verifier's outputs on the Hadamard round is computationally indistinguishable from $M_{XZ}(\rho, h)$.
\iffalse    
    Suppose that for all $\lambda\in\bbN$ and $h\in\zo^*$ \Ethan{or $\zo^n$?},
    $\PMeasureStar$ passes the testing round with probability $1-\negl(\lambda)$.
    Then, under the QLWE assumption, there exists some $\rho$ so that for all $h$,
    The verifier's output on the Hadamard round is $\negl(\lambda)$-computationally indistinguishable from $M_{XZ}(\rho, h)$.
\fi
\end{lemma}

@@

There are several problems with using $\PiNaive$ as a $\SampBQP$ protocol. first, since the verifier doesn't get a sample if he had chosen the testing round in Step~\ref{step:c-urmila-in-naive}, so the protocol has  completeness error at least $1/2$. And since Protocol~\ref{proto:urmila4} does not check anything on the Hadamard round, a cheating prover can give up passing the testing round and breaks the commitment on the Hadamard round. Such cheating prover only has a constant $1/2$ probability of being caught, resulting in constant soundness error.

% Note that in $\PiNaive$ the verifier only learns the output in the Hadamard round.
% So $\PiNaive$ is not a $\QPIP_0$ with good completeness or soundness.

@@

We now introduce our $\QPIP_0$ protocol $\PiSampZ$ for $\SampBQP$.
It is essentially a $m$-fold parallel repetition of $\PiNaive$.
Instead of having each copy running testing round or Hadamard round uniformly at random, we randomly pick one copy to run Hadamard round to get our samples, and run testing round on all other $m-1$ copies.
In the description of our protocol below, we describe $\PiNaive$ and $\PiMeasure$ in details  in order to introduce notations that we need in our analysis.

@@

Inputs:
    \begin{itemize}
        \item Security parameter $1^\lambda$ for $\lambda\in\bbN$.
        \item Accuracy parameter $1^{1/\eps}$ for the $\SampBQP$ instance
        \item Input $x\in\zo^{\poly(\lambda)}$ for the $\SampBQP$ instance
    \end{itemize}

Ingredient: Let $m=O(1/\eps^2)$ be the number of parallel repetitions to run.

@@

The intuition behind this theorem is that we send out $m$ independent copies of the naive $\QPIP_0$ and do testing on randomly chosen $m-1$ copies of them, and do Hadamard round and calculate results from the remaining one copy. Therefore if the prover wants to cheat, i.e. send something not binding in the Hadamard round copy, he would be caught with probability $1-1/m$. However, the prover might create correlations between copies that let him cheat with probability much higher than $1/m$. This is a common problem in trying to create parallel repetition protocols. We are able to bound the soundness error as $O(1/\sqrt{m})$. To prove that, we use the following partition lemma from \cite{parallelrep}, which intuitively says that for each $i\in[m]$, there exist two efficient ``projectors" \footnote{Actually they are not projectors, but for the simplicity of this discussion let's assume they are.} $G_{0,i}$ and $G_{1,i}$ in the prover's internal space with $G_{0,i}+G_{1,i} \approx Id$.  $G_{0,i}$ represents the subspace that are rejected with high probability in the testing round of $i$-th copy; $G_{1,i}$  represents the subspace accepted with ``high probability"\footnote{This is also not technically correct as we will discuss later.} in the testing round of $i$-th copy.  $G_{0,i}$ and $G_{1,i}$ need to be efficient because $\PiMeasure$ only gives computational security. Once we have $G_{0,i}$ and $G_{1,i}$, we can split up the prover's internal state $\ket{\psi}$ like conditional probabilistic events.  We  split up  $\ket{\psi}$ into $G_{0,1}\ket{\psi}+ G_{0,2}G_{1,1}\ket{\psi} + G_{0,3}G_{1,2}G_{1,1}\ket{\psi} + \dots +  G_{0,m} G_{1,m-1}\dots G_{1,2}G_{1,1}\ket{\psi} + G_{1,m} G_{1,m-1}\dots G_{1,2}G_{1,1}\ket{\psi}$, where for example, $G_{0,2}G_{1,1}\ket{\psi}$ intuitively corresponds to passing the testing round of the first copy then getting rejected on the testing round of the second copy. Since each of the  substates in  $\{G_{0,1}\ket{\psi}, G_{0,2}G_{1,1}\ket{\psi} , G_{0,3}G_{1,2}G_{1,1}\ket{\psi} , \dots ,  G_{0,m} G_{1,m-1}\dots G_{1,2}G_{1,1}\ket{\psi} \}$ all have  probability $1-1/m$ of being caught by one of the $m-1$ testing rounds, they have well-bounded
 probability of being accepted in Step~\ref{step:multi-testing} of Protocol~\ref{proto:QPIP0samp}, so the output in  Step~\ref{step:multi-testing} would mostly corresponds to $G_{1,m} G_{1,m-1}\dots G_{1,2}G_{1,1}\ket{\psi}$. However, since $G_{1,i}$ don't commute with $G_{1,j}$, $G_{1,m} G_{1,m-1}\dots G_{1,2}G_{1,1}\ket{\psi}$ is only binding for the $m$-th copy, so this simplified strategy is not quite correct. We were able to fix this issue with a more careful split up of $\ket{\psi}$. Also as a careful reader might have noticed, the prover's space doesn't happen to split nicely into parts of very high accept probability in the testing round and parts with very low accept probability, as there will always be something in the middle. In \cite{parallelrep} this is solved by doing eigenvalue estimation to calculate the accept probability, then split the space into parts that are accepted with probability higher or lower than a small threshold $\gamma$. However, states with accept probability really close to the threshold $\gamma$ can not be classified, so we need to average over randomly chosen $\gamma$ to have $G_{0,i}+G_{1,i} \approx Id$.

 %but there will always be some states  therefore we introduce  an extra parameter $\gamma$ that corresponds to the probability cut-off line that to distinguish the accepted and rejected part, and we need to average over $\gamma$.

@@

%For example, $G_{0,3} G_{1,5} \ket{\psi}$ corresponds to the events of passing the testing round on the fifth copy then fails the testing round on the third copy.

@@

where $P_{i,pk_i,y_i,acc}$ are projector to the states that $i$-th testing round accepts with $pk_i,y_i$, including the last measurement the prover did before sending $\vec{a}$.  This means that $\ket{\psi_{0,i,\gamma}}$ is rejected by the $i$-th testing round with high probability.

@@

% \begin{align*}  
%   \Pr\left[M_{\regX_i}\circ \ext_i\left(\frac{\ket{\{c\}}_{\regC}\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}}{\|\ket{\psi_1}_{\regX,\regZ}\|}\right)\in \Acc_{pk_i,y_i}\right]=1-\negl(\secpar).
%   \end{align*}
This will imply that    $\ket{\psi_{1,i,\gamma}}$ is binding to the $i$-th Hadamard round.

@@

where we abuse the notation and use $\ket{\psi_{err,i,\gamma}}$ to denote the error state we get from decomposing $\ket{\psi_{1^{i-1},\gamma}}$. %\hannote{Note that the prover's state is actually a mixed state. In particular, $\ket{\psi}$ depends on $pk$ and we are considering the mixed state averaged over the verifier's choice of $pk$.}

@@

We now proceed to prove that $z_{good}$ is close to the ideal distribution.

@@

The prover tries to run $U_0$ by taking the verifier's input as $i$-th copy of Protocol~\ref{proto:urmila4} in Protocol~\ref{proto:QPIP0samp} and simulating other $m-1$ copies by himself. The prover then picks a uniformly random $\gamma$ and  tries to generate $\ket{\psi_{1^{i-1}1,\gamma}}$ by applying $G_{i,1,\gamma}G_{i-1,1,\gamma} \cdots G_{2,1,\gamma}G_{1,1,\gamma}$. If the prover fails to generate $\ket{\psi_{1^{i-1}1,\gamma}}$, he throws out everything and aborts by sending $\bot$ back.   On the fourth round,  If it's a testing round the prover reply with the $i$-th register of $\ext_i\left(\frac{\ket{e_j}_{\regC}\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}}{\|\ket{\psi_1}_{\regX,\regZ}\|}\right)$, where $\ext_i$ is specified in property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}. If it's the Hadamard round  the prover  runs $U$ and checks whether every copy except the $i$-th copy would be accepted. If all $m-1$ copies are accepted, he replies with the $i$-th copy, otherwise reply $\bot$.

@@

By  property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}, the above strategy is accepted with probability $1-\negl(n)$ when the prover didn't abort. Since the prover's strategy is also efficient, by Lemma~\ref{lem:naive-qpip0-binding},
\begin{align}
    (d,z_{composite,i})\approx_{c, O(\eps)}(d,z_{ideal,i})
\end{align}
